---
title: "HW6"
output: 
 html_notebook:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  markdown: 
    wrap: sentence
---

Use this R Notebook document to answer the questions and document your work.
Enter the R code used to answer each question in the corresponding R code chunk.
Write any textual explanations **outside** of the chunks.
When we grade, we will *hide* the code and just look at the R output and your written answers.
Attempt to clean up your code as much as possible so that only the necessary lines remain.

When you are done:

1.  Select '**Restart R and Run All Chunks**' from the 'Run' dropdown menue.
2.  Save (File -\> Save)
3.  Click 'Preview' to bring up the `HW6.nb.html` file. Check through this to make sure it rendered correctly.
4.  Upload the `HW6.nb.html` to Canvas.

Necessary libraries

```{r echo = F,message=FALSE, warning=FALSE}
# Necessary packages 
library(ggplot2) 
library(emmeans)
library(lme4)
library(lmerTest)
library(PLS205)
```

## Question 1

A researcher was interested in examining the effect of pasture species composition on beta-carotene content in eggs of pastured poultry.
She evenly divided each of four fields into five sections and overseeded each section with one of the following: 1) annual rye (grass); 2) fescue (grass); 3) white clover (legume); 4) vetch (legume); or, 5) clover/fescue mix.
A separate flock of 10 birds was put out in each section.
After one-week, eggs from each flock were collected daily and combined for determining beta-carotene content of the yolks.
These values in $\mu_g/g$ were averaged after 10 days.

```{r echo=FALSE}
data_1 = read.csv('egg_carotene.csv')
data_1$Section = as.factor(data_1$Section)
str(data_1)
```

### 1.1 Describe the design of this experiment in detail.

**Design**: **Response**:

| Structure | Variable      | Type        | \# levels | Blocks | EU      |
|-----------|---------------|-------------|-----------|--------|---------|
| Treatment | Species       | Categorical | 5         | Block  | Section |
| Design    | Block         | Categorical | 4         |        |         |
|           | Species:Block | Categorical | 20        |        |         |
|           | Section       | Categorical | 20        |        |         |
| Response  | Beta          | Numerical   | 20        |        |         |

> ENTER YOUR RESPONSE HERE

### 1.2 Assess whether the assumptions of a linear model are adequately satisfied.

Regardless of what you find, do not do a transformation for the rest of the questions.

```{r}

```

> ENTER YOUR RESPONSE HERE

### 1.3 Estimate the differences between all pairs of pasture covers. Include appropriate units based on your choice of transformation (or no transformation)

```{r}
```

> ENTER YOUR RESPONSE HERE

### 1.4 Is there any evidence of an interaction between Field and Species cover? ie. do differences among covers change across the fields?

Is the interaction replicated?
If not, you can't answer this question with an ANOVA, but you can tentatively assess with an interaction plot.
Describe whether it looks like there are large interactions present.

```{r}
```

> ENTER YOUR RESPONSE HERE

### 1.5 Did the researcher gain information by utilizing a RCBD relative to a CRD where the sections of all fields were randomized together?

```{r}
```

> ENTER YOUR RESPONSE HERE

### 1.6 Instead of randomizing all 5 plots on the 4 fields together, speculate on the efficiency had she used only one field.

In this case, to maintain the same replication, she would divide it into 20 smaller plots, and use only 2 birds per plot (instead of 10).
Speculate on how the MSE/$$s^2$$ would compare to the original RCBD, or the CRD over the four fields.
**Is there any advantage to spreading the experiment over 4 fields** Is there any disadvantage?

> ENTER YOUR RESPONSE HERE

### 1.7 Add 1 $\mu g/g$ to each value in Field F4. Does this produce any effect on the SS_Species, MS_Species, or F-value? Explain why or why not in one sentence.

In this case (with Field F4 being much more productive), did the importance of Blocking increase or decrease?

The following code makes a modified dataset with the increase of 1 in the Beta values for Field F4

```{r}
data_2_modified = data_1
data_2_modified$Beta[data_2_modified$Field == 'F4'] = data_2_modified$Beta[data_2_modified$Field == 'F4'] + 1
```

> ENTER YOUR RESPONSE HERE
