---
title: "HW2"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
editor_options: 
  markdown: 
    wrap: sentence
---

Use this R Notebook document to answer the questions and document your work.
Enter the R code used to answer each question in the corresponding R code chunk.
Write any textual explanations **outside** of the chunks.
When we grade, we will *hide* the code and just look at the R output and your written answers.
Attempt to clean up your code as much as possible so that only the necessary lines remain.

When you are done:

1.  Select 'Run All' from the 'Run' dropdown menu.
2.  Save (File -\> Save)
3.  Click 'Preview' to bring up the `HW2.nb.html` file. Check through this to make sure it rendered correctly.
4.  Upload the files: `HW2.nb.html` and `HW2.Rmd` to Canvas.

------------------------------------------------------------------------
# Part 1

```{r}
library(PLS205)
library(emmeans)
```

## Experiment

A researcher is interested in the difference in root growth between two wheat varieties.
She plants 16 seeds of each variety in a completely randomized design and harvests the roots after 1 week.
She records the root length (cm) of each plant.

```{r echo = FALSE}
# This loads the data for this question
data_1 <- read.csv('wheat_roots.csv')
data_1$Variety = as.factor(data_1$Variety)
# This prints a summary of the data table
str(data_1)
```

### Describe the design of this experiment in detail.


Use the following table.
You can also fill out the table in an excel document, and then paste it into [this website](http://www.tablesgenerator.com/markdown_tables) to generate the Markdown table.

**Level of Replication**: Treatment levels

| Structure | Variable | Type | \# levels | Replicate | Experimental Unit |
|-----------|----------|------|-----------|-----------|-------------------|
| Treatment |Variety   | Factor| 2         | None      | Plant             |
| Design    |Plant    | Factor | 32          |           |                   |
| Response  | Root_length | Numeric     | 32          |           |                   |

### Do any columns of `data_1` need to be changed to be consistent with the design table?

Yes. Our plant numbers, as they are descriptive/categorical, should be formatted as `chr` or `fctr`. I'm going to use `chr` since they are all unique and do not need a finite number of categorical values.

Make any necessary changes here:

```{r}
data_1$Plant = as.character(data_1$Plant)
str(data_1)
```

> ENTER YOUR RESPONSE HERE

### Write the linear model for this experiment using the R function `lm()` or `lmer()` as appropriate:

Explain why you chose `lm()` or `lmer()`

```{r}
rootmod = lm(Root_length ~ Variety, data=data_1)
rootmod
```

> ENTER YOUR RESPONSE HERE

I chose `lm()` given that there were no random variables in this experiment.

### Make the model diagnostic plots shown in lab. Is there reason to be concerned about any of the model assumptions?

```{r}
# enter your code here
pls205_diagnostics(rootmod)
```

> ENTER YOUR RESPONSE HERE

The normality looks good, but there's a possibility that the variance for Variety B may be less than A, but it doesn't look egregious, and I would argue that our assumptions are fine.

### User the `emmeans()` and `contrast()` functions to estimate the difference in root lengths between the two varieties

Report a 90% confidence interval for the difference

```{r}
# enter your code here
means_model <- emmeans(rootmod, spec = 'Variety')
diffmod <- contrast(means_model,'pairwise')
meanssummary = summary(means_model, level = 0.9, infer = c(T,T))
diffsummary = summary(diffmod, level = 0.9, infer = c(T,T))
print(diffsummary)
```

> ENTER YOUR RESPONSE HERE

Difference in root length is around 3.47, with a CI of 2.08 to 4.86.

## Repeat this analysis "by hand"

The next few questions ask you to repeat the analysis withing using the modeling functions `lm()` and `emmeans` to show that you understand the mechanisms underlying the analysis

To help you analyze this data, I've separated out the data from each variety into a vector:

```{r}
Variety_A <- data_1$Root_length[data_1$Variety == 'A']
Variety_B <- data_1$Root_length[data_1$Variety == 'B']
```

Use the variables `Variety_A` and `Variety_B` below for your calculations.

For the following 4 questions, you can check your answers below with the `t.test` function, but you must write out the calculations by hand.
You can use functions like `mean()` and `sd()`, `qt`, etc.

### Estimate the difference in root lengths between the two varieties

```{r}
#This should be the difference in mean root length, I would think.
differ = abs(mean(Variety_A) - mean(Variety_B))
differ
```

> ENTER YOUR RESPONSE HERE

The difference appears to be 3.47 (which matches, hooray!)

### Estimate the Standard Error of this difference. Assume the variances of the two varieties are the same, and find a pooled estimate of the variance.

```{r}
#So if we wanna pool all the deviations together, we can use the pooled SED. Using 'more accurate' form from lecture
sed = sqrt((var(Variety_A) / 16) + (var(Variety_B) / 16))
sed

```

> ENTER YOUR RESPONSE HERE

Standard error is 0.817

### Form a 90% confidence interval for this difference

```{r}
#Going to calculate tcrit with alpha = 0.1 (for the 90% interval), then multiply that by the standard error and add/subtract it to the estimated difference.
alpha = 0.1
tcrit = qt(alpha/2, df = 30, lower.tail = F)
cilow = differ - (sed * tcrit)
cihi = differ + (sed * tcrit)

tcrit
cilow
cihi
```
T-sub-c is 1.70, and the Confidence Interval runs from 2.08 to 4.86