---
title: "HW5"
output: 
 html_notebook:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: inline
---

**Note:** Before submitting your document, re-run all your R code chunks
by clicking the down-arrow next to "Run" in the document's menu bar and
selecting "**Restart R and Run All Chunks**". Then click "Preview" and
go through each answer to make sure the R code and any associated output
is provided.

Necessary libraries

```{r echo = F,message=FALSE}
library(ggplot2)
library(emmeans)
library(lme4)
library(lmerTest)
library(PLS205)
library(multcomp)
```

An experiment was run to examine the combined effects of Exercise
regimes on weight loss in dogs. They hypothesized that this may depend
on diet, so used a factorial experimental design. The experiment used
four commercial dog foods (Food_A - Food_D) and three exercise regimes
(WALK: daily walking, LOW: 30 min run 2x/week, HIGH: 30 min run 6 days
per week). They recruited 96 dogs from a nearby vet clinic and assigned
each to one of the 4 food brands and one of the 3 exercise regimes.
Weights were collected at the beginning of the experiment and after 3
months. Responses were recorded as % change in weight after 3 months for
each dog.

Here are the data:

```{r}
dog_data = read.csv('dog_data.csv')
str(dog_data)
```

### 1.1

**Create the Design table for this experiment. Use `Exercise` as the
focal treatment. Explain your choices for EU and Replicate.**

| Structure     | Variable | Type | #levels | Replicate | EU  |
|---------------|----------|------|---------|-----------|-----|
| Focal Trt     |          |      |         |           |     |
| Moderator Trt |          |      |         |           |     |
| Combo Trt     |          |      |         |           |     |
| Design        |          |      |         |           |     |
| Response      |          |      |         |           |     |

> ENTER YOUR RESPONSE HERE

### 1.2

**Does the data.frame need to be updated to match the design table? If
so, make necessary changes:**

```{r}

```

### 1.3

**Make a plot of the data, separating by focal and moderator variables
as done in lab.**

**1) Does it look like the different exercise regimes affected weight
loss differently?** Is this a question about main, specific, or
interaction effects?

**2) Does the differences among exercise regimes seem to change between
the foods?** Is this a question about main, specific, or interaction
effects?

```{r}

```

> ENTER YOUR RESPONSE HERE

### 1.4

Write a linear model to estimate **specific effects** of exercise at
each level of food.

```{r}

```

### 1.5

**Compare the 3 exercise regimes for each type of food. Use
`contrasts()` to make a table of `trt.vs.ctrl` comparisons for each type
of food, using `WALK` as the control regime. Produce confidence
intervals such that the chance that ANY of the confidence intervals do
not include the true value is less than 5%.**

**Can you conclude that there are different effects of any of these 3
exercise regimes on weight gain using any of the 4 foods? Use**
$\alpha = 0.05$**.**

```{r}

```

> ENTER YOUR RESPONSE HERE

### 1.6

**Answer the similar question using an ANOVA. Can you conclude that
there are different effects among any of these 3 exercise regimes on
weight gain using any of the 4 foods? Use** $\alpha = 0.05$**. What is
the difference between this question and the one in 1.5?**

```{r}

```

> ENTER YOUR RESPONSE HERE

### 1.7

**Did effects of changing to either HIGH or LOW exercise change among
diets?**

**Calculate interaction effects for the HIGH-WALK and LOW-WALK contrasts
across foods. Use `pairwise` contrasts for each contrast class among
foods, and adjust your confidence intervals for the number of contrast
classes.** **Use** $\alpha = 0.05$.

**Use the CLD function to group the contrasts for each food into groups
that cannot be distinguished. Again, adjust the confidence level fo the
number of contrast classes.**

Write a complete statement describing a significant interaction contrast
that you observe.

```{r}

```

> ENTER YOUR RESPONSE HERE

### 1.8

**Repeat the last question using ANOVA. Compare the minimum p-value for
the pairwise interaction contrasts and the p-value from the ANOVA.**

```{r}

```

> ENTER YOUR RESPONSE HERE
