---
title: "HW 8"
output: 
 html_notebook:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  markdown: 
    wrap: sentence
---

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(lme4)
library(lmerTest)
library(emmeans)
library(ggplot2)
library(PLS205)
library(multcomp)
```

A split-plot experiment for studying the effect of temperature on corrosion of steel bars treated with four coatings (C1, C2, C3, C4).
They used three furnace temperatures (360C, 370C, 380C).
It takes considerable time to reset the furnace and reach a new equilibrium temperature.
Therefore, each the furnace was set twice to each temperature (in a randomized order), and then while it was at the set temperature, four bars, one with each coating, were positioned in the furnace at randomized locations and heated.
Each bar was then tested individually for corrosion resistance.

Data from the experiment are provided here:

```{r}
corrosion = read.csv('corrosion.csv',stringsAsFactors = T)
corrosion$Bar = as.factor(corrosion$Bar)
corrosion$Trial = as.factor(corrosion$Trial)
str(corrosion)
```

## 1)

**Create the design table for this experiment**

**Correct any of the variables in the data.frame to correspond to the design table**

**Use `Temp` as the focal treatment.**

| Structure | Variable                       | Type        | #levels | Block | EU    |
|------------|------------|------------|------------|------------|------------|
| Focal     | Temp                           | Categorical | 3       | Coat  | Trial |
| Moderator | Coat                           | Categorical | 4       | None  | Bar   |
| Combo     | Temp:Coat                      | Categorical | 12      | Trial | Bar   |
| Design    | go look at class 17 and 18 pdf |             |         |       |       |
|           |                                |             |         |       |       |
|           |                                |             |         |       |       |
|           |                                |             |         |       |       |
| Response  |                                |             |         |       |       |

```{r}
```

> [9 points]

> ENTER YOUR RESPONSE HERE

## 2)

**Write a linear model for these data**

```{r}


```

> [5 points]

> ENTER YOUR RESPONSE HERE

## 3)

**Run appropriate model diagnostics**

**Be sure to do diagnostics on all EU. Do you see any issues? *Do not do any transformation in either case***

```{r}


```

> [6 points]

> ENTER YOUR RESPONSE HERE

## 4)

**How strong is the evidence that the Temperature has any effect on corrosion?**

```{r}


```

> [6 points]
>
> ENTER YOUR RESPONSE HERE

## 5)

**Can you conclude that the coating modifies the effect of temperature?**

**What contrast among Coatings appears to change the most among temperatures? Include a figure demonstrating your results**

```{r}


```

> [6 points]

> ENTER YOUR RESPONSE HERE

## 6)

**Are the estimated standard errors of the specific effects of temperature larger or smaller than the interaction effects (ie the effects of coating on temperature effects)? Is this unusual for a factorial? Why?**

> [6 points]

> ENTER YOUR RESPONSE HERE

## 7)

**Speculate on whether the experiment would have been more or less precise if it had been run as a completely randomized design instead of a split-plot.**

**In this case, only a single bar would have been heated in an oven at a time. How much longer would the experiment have taken?**

> [6 points]

> ENTER YOUR RESPONSE HERE

## 8)

**Could you report the main effect of `Coat`** **and/or `Temp` in this design? Why or why not?**

**You do not need to do this analysis, but explain how it could be done, or why it could not.**

> [ 6 points ]

> ENTER YOUR RESPONSE HERE
